---
title: "California Wildfires Report"
author: "Abigail Chutnik, Ben Lewis, Hangliang Ren, Alex Andrei Ungureanasu"
date: "11/20/2021"
output: html_document
---

```{r setup, include=FALSE}
source("analysis.R")
source("table.R")
source("charts.R")
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

The western United States has faced increasing amounts of wildfires in the last 
decade. Wildfires are a major threat to the population's own health and the
health of our environment. Currently, many speculate that the increasing amounts
of wildfires may be due to climate change and it's associated increasing
temperatures. To understand more about wildfires in the West coast,
our report investigates the trends in wildfires, specifically in California. 

## Summary

In order to analyze the trends within the last decade, our analysis focuses on
the years 2013 to 2020. Wildfires show tremendous consequences on our
environment as each year, **1,070,651** acres of land were impacted by wildfires.
Specifically, Riverside county suffered from wildfires the most during this
time period. Additionally, **2017** was the worst year for wildfires in
California, since the most occurred during that year. We also wanted to look
at the proportion of equipment used by firefighters and personnel, and found
that engines are used the most. The proportions of these
equipment used are helicopters, dozers, and engines and their
proportions are, respectively, **0.017**, **0.075**, and **0.117**. Lastly,
we were curious about the average labor participated in handling the
wildfires, which were about **580** in 2013 and decreased to **227** in
2017, with the average change of **60%**. 


## Table

This table is included to showcase the trends in wildfire deaths over the years.
In this table, we noticed an increase in wildfire related deaths from
2017 to 2018, and a sudden decrease in 2019.

```{r echo=FALSE}
table
```


## Chart 1

This bar graph includes the number of acres burned per year in order to
showcase the amount of damage done by wildfires over time.

```{r echo=FALSE}
gg_bar
```

## Chart 2

This line graph demonstrates the frequencies of equipment and machinery used
by firefighters and personnel to handle wildfires over time. Trends for these
types of equipment are quite similar, besides the magnitude of each as shown
with the high amount of "Engines" being used.

```{r echo=FALSE}
gg_line
```


## Chart 3

```{r echo=FALSE}
gg_ca
```

This map shows the instance of wildfires in California through the last decade
in order to highlight areas where there is a higher abundance of wildfires.
